Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-27 paper/paper.tex
Change Figure to Fig.

The DPDS algorithm guarantees that tasks from lower priority workflows are
always deferred when higher-priority tasks are available, but lower-priority
tasks can still occupy idle VMs when higher-priority tasks are not available.
However, because there is no preemption, long-running low-priority tasks may
delay the execution of higher-priority tasks. In addition, tasks from low
priority workflows may be executed even though there is no chance that those
workflows will be completed within the current budget and deadline.
Figure~\ref{fig:algorithm-example}a shows an example schedule generated using
the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
backfill idle VMs when tasks from higher priority workflows are not available.

@@

An example schedule generated by SPSS (Figure~\ref{fig:algorithm-example}b)
shows how SPSS tends to start many workflows in parallel, running each workflow
over a longer period of time on only a few VMs to minimize cost. In comparison,
the dynamic algorithms tend to run one workflow at a time across many VMs in
parallel.

@@

Pareto ensembles contain a small number of larger workflows and a large number
of smaller workflows. Their sizes  
%in a Pareto ensemble 
are chosen according to a Pareto distribution. The distribution was modified so
that the number of large workflows (of size $\geq$ 900) is increased by a small amount to
produce a ``heavy-tail''. This causes Pareto ensembles to have a slightly larger
number of large workflows, which reflects behavior commonly observed in many
computational workloads. 
%An example of the distribution of workflow sizes that
%occurs in a Pareto ensemble is shown in Figure~\ref{fig:ensemble-pareto}.

@@

Figure~\ref{fig:distributions} shows the percentage of simulations for which
each algorithm achieved the highest score for a given ensemble type. This
experiment was conducted using all five applications, with all five types of
ensembles. For each application and ensemble type, 10 random ensembles of 50
workflows each were created. Each ensemble was simulated with all three
algorithms using 10 budgets and 10 deadlines (1000 simulations per
application, ensemble type, and algorithm). The best scores percentage is
computed by counting the number of times that a given algorithm achieved the
highest score and dividing by 1000. Note that it is possible for multiple
algorithms to get the same high score (to tie), so the numbers do not
necessarily add up to 100\%. The sum is much higher than 100\% in cases where
the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which
are very similar algorithms, often get the same high score.

@@

There are several interesting things to notice about
Figure~\ref{fig:distributions}. The first is that, in most cases, SPSS
significantly outperforms both dynamic algorithms (DPDS and WA-DPDS). This is
attributed to the fact that SPSS is able to make more intelligent scheduling
and provisioning decisions because it has the opportunity to compare different
options and choose the one that results in the best outcome. In comparison,
the dynamic algorithms are online algorithms and are not able to project into
the future to weigh the outcomes of their choices.

@@

Figure~\ref{fig:stretching} shows the relative performance of the algorithms
as the scaling factor is increased from 1 to 16. The figure only shows results
for the CyberShake application with pareto unsorted ensembles, but similar
results were found for other ensemble types for both Montage and CyberShake.
Each data point in the figure represents 500 simulations (5 ensembles with 10
budgets and 10 deadlines). The best scores percent was calculated the same way
as it was for Figure~\ref{fig:distributions}.

@@

Figure~\ref{fig:variances} shows the results of simulations for estimate
errors ranging from $0\%$ to $50\%$. These simulations were done for 10
ensembles of 50 workflows using 10 budgets and 10 deadlines for each error
value and algorithm (3000 simulations per error value). The figure shows box
plots of the ratio of actual cost to budget and actual makespan to deadline.
The actual cost and makespan are the simulated cost and makespan of the
ensemble. The ratio of actual values to constraints indicates whether the
actual value exeeded the constraint: Values less than 1 indicate that the
constraint was not exceeded, and values greater than 1 indicate that the
constraint was exceeded.

Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
figure illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions before execution and is not able to adapt
to changing circumstances.

@@

Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
interesting thing to notice about this figure is that, for all three
algorithms, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of
the fact that they can adapt to poor estimates. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a
particularly clever optimization. The SPSS algorithm tends to schedule
workflows early, using up the budget long before the deadline is reached. This
is a consequence of the deadline distribution function in SPSS, which prevents
workflows from starting late. This is shown by the gantt chart in
Figure~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
up workflows at the beginning of the timeline. In comparison, the dynamic
algorithms tend to spread out workflow starts over the entire duration as
shown in Figure~\ref{fig:algorithm-example}.a. The consequence of this
behavior is that SPSS plans tend to use up the budget but leave plenty of time
before the deadline. So when the runtime of the plan is increased by
introducing errors, the SPSS plan has some room to expand without exceeding
the deadline.

@@

Figure~\ref{fig:delays} shows the ratios of actual values to constraints when
the provisioning delay is increased from 0 seconds up to 15 minutes. These
simulations were done using 10 uniform unsorted ensembles of 50 workflows from
the Montage application, using 10 deadlines (1000 simulations per algorithm
and provisioning delay). As in Figure~\ref{fig:variances}, the y-axis in each
plot represents the ratio of the actual simulated value to the constraint
value for the budget constraint (actual cost) or the deadline constraint
(actual makespan). These ratios indicate how frequently and by how much the
actual simulated values exceeded (or did not exceed) the constraints.

@@

The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not. The explanation for this behavior is also similar. The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget. Figure~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget. In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget. With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.

% Figure~\ref{fig:delays_scores} shows the relative performance as a
% percentage of the number of high scores achieved by each algorithm as the
% provisioning delay increases. This figure shows that, as the delay
% increases, the relative performance of the static algorithm increases as
% well. This behavior is a result of the fact that the static algorithm is, in
% essence, cheating by using more time and money than the dynamic algorithms.
% In the previous section we showed how the relative performance of the
% algorithms remains the same when a uniform error is applied to the runtime
% constraints. In that case, the dynamic algorithms adapted to the error by
% rearranging the schedule to accomplish the same amount of work given the
% same deadline and budget. In this case, the dynamic algorithms adapt by
% performing less work (executing fewer workflows) to remain within the
% constraints while accounting for the delays. In both cases, the static
% algorithm completed the same amount of work, but did so by exceeding the
% constraints.

@@

Figure~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and 800 tasks. The ensembles were generated using a constant distribution equal to the workflow size desired. Two different applications were used: SIPHT and CyberShake. Each box summarizes the results of 2 applications, 10 random seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.

Figure~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes. For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10 minutes. Considering that the largest workflows used in this experiment are still relatively small (maximum of 800 tasks), and that real workflows are often much larger (workflows with tens of thousands of tasks are common, and workflows with hundreds of thousands or millions of tasks are not impossible), it is unlikely that SPSS will be practical for ensembles of very large workflows.

>>>>

The DPDS algorithm guarantees that tasks from lower priority workflows are
always deferred when higher-priority tasks are available, but lower-priority
tasks can still occupy idle VMs when higher-priority tasks are not available.
However, because there is no preemption, long-running low-priority tasks may
delay the execution of higher-priority tasks. In addition, tasks from low
priority workflows may be executed even though there is no chance that those
workflows will be completed within the current budget and deadline.
Fig.~\ref{fig:algorithm-example}a shows an example schedule generated using
the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
backfill idle VMs when tasks from higher priority workflows are not available.

@@

An example schedule generated by SPSS (Fig.~\ref{fig:algorithm-example}b)
shows how SPSS tends to start many workflows in parallel, running each workflow
over a longer period of time on only a few VMs to minimize cost. In comparison,
the dynamic algorithms tend to run one workflow at a time across many VMs in
parallel.

@@

Pareto ensembles contain a small number of larger workflows and a large number
of smaller workflows. Their sizes  
%in a Pareto ensemble 
are chosen according to a Pareto distribution. The distribution was modified so
that the number of large workflows (of size $\geq$ 900) is increased by a small amount to
produce a ``heavy-tail''. This causes Pareto ensembles to have a slightly larger
number of large workflows, which reflects behavior commonly observed in many
computational workloads. 
%An example of the distribution of workflow sizes that
%occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.

@@

Fig.~\ref{fig:distributions} shows the percentage of simulations for which
each algorithm achieved the highest score for a given ensemble type. This
experiment was conducted using all five applications, with all five types of
ensembles. For each application and ensemble type, 10 random ensembles of 50
workflows each were created. Each ensemble was simulated with all three
algorithms using 10 budgets and 10 deadlines (1000 simulations per
application, ensemble type, and algorithm). The best scores percentage is
computed by counting the number of times that a given algorithm achieved the
highest score and dividing by 1000. Note that it is possible for multiple
algorithms to get the same high score (to tie), so the numbers do not
necessarily add up to 100\%. The sum is much higher than 100\% in cases where
the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which
are very similar algorithms, often get the same high score.

@@

There are several interesting things to notice about
Fig.~\ref{fig:distributions}. The first is that, in most cases, SPSS
significantly outperforms both dynamic algorithms (DPDS and WA-DPDS). This is
attributed to the fact that SPSS is able to make more intelligent scheduling
and provisioning decisions because it has the opportunity to compare different
options and choose the one that results in the best outcome. In comparison,
the dynamic algorithms are online algorithms and are not able to project into
the future to weigh the outcomes of their choices.

@@

Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
as the scaling factor is increased from 1 to 16. The figure only shows results
for the CyberShake application with pareto unsorted ensembles, but similar
results were found for other ensemble types for both Montage and CyberShake.
Each data point in the figure represents 500 simulations (5 ensembles with 10
budgets and 10 deadlines). The best scores percent was calculated the same way
as it was for Fig.~\ref{fig:distributions}.

@@

Fig.~\ref{fig:variances} shows the results of simulations for estimate
errors ranging from $0\%$ to $50\%$. These simulations were done for 10
ensembles of 50 workflows using 10 budgets and 10 deadlines for each error
value and algorithm (3000 simulations per error value). The figure shows box
plots of the ratio of actual cost to budget and actual makespan to deadline.
The actual cost and makespan are the simulated cost and makespan of the
ensemble. The ratio of actual values to constraints indicates whether the
actual value exeeded the constraint: Values less than 1 indicate that the
constraint was not exceeded, and values greater than 1 indicate that the
constraint was exceeded.

Fig.~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
figure illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions before execution and is not able to adapt
to changing circumstances.

@@

Fig.~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
interesting thing to notice about this figure is that, for all three
algorithms, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of
the fact that they can adapt to poor estimates. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a
particularly clever optimization. The SPSS algorithm tends to schedule
workflows early, using up the budget long before the deadline is reached. This
is a consequence of the deadline distribution function in SPSS, which prevents
workflows from starting late. This is shown by the gantt chart in
Fig.~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
up workflows at the beginning of the timeline. In comparison, the dynamic
algorithms tend to spread out workflow starts over the entire duration as
shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
behavior is that SPSS plans tend to use up the budget but leave plenty of time
before the deadline. So when the runtime of the plan is increased by
introducing errors, the SPSS plan has some room to expand without exceeding
the deadline.

@@

Fig.~\ref{fig:delays} shows the ratios of actual values to constraints when
the provisioning delay is increased from 0 seconds up to 15 minutes. These
simulations were done using 10 uniform unsorted ensembles of 50 workflows from
the Montage application, using 10 deadlines (1000 simulations per algorithm
and provisioning delay). As in Fig.~\ref{fig:variances}, the y-axis in each
plot represents the ratio of the actual simulated value to the constraint
value for the budget constraint (actual cost) or the deadline constraint
(actual makespan). These ratios indicate how frequently and by how much the
actual simulated values exceeded (or did not exceed) the constraints.

@@

The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not. The explanation for this behavior is also similar. The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget. Fig.~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget. In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget. With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.

% Fig.~\ref{fig:delays_scores} shows the relative performance as a
% percentage of the number of high scores achieved by each algorithm as the
% provisioning delay increases. This figure shows that, as the delay
% increases, the relative performance of the static algorithm increases as
% well. This behavior is a result of the fact that the static algorithm is, in
% essence, cheating by using more time and money than the dynamic algorithms.
% In the previous section we showed how the relative performance of the
% algorithms remains the same when a uniform error is applied to the runtime
% constraints. In that case, the dynamic algorithms adapted to the error by
% rearranging the schedule to accomplish the same amount of work given the
% same deadline and budget. In this case, the dynamic algorithms adapt by
% performing less work (executing fewer workflows) to remain within the
% constraints while accounting for the delays. In both cases, the static
% algorithm completed the same amount of work, but did so by exceeding the
% constraints.

@@

Fig.~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and 800 tasks. The ensembles were generated using a constant distribution equal to the workflow size desired. Two different applications were used: SIPHT and CyberShake. Each box summarizes the results of 2 applications, 10 random seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.

Fig.~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes. For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10 minutes. Considering that the largest workflows used in this experiment are still relatively small (maximum of 800 tasks), and that real workflows are often much larger (workflows with tens of thousands of tasks are common, and workflows with hundreds of thousands or millions of tasks are not impossible), it is unlikely that SPSS will be practical for ensembles of very large workflows.
